### 2021-10-17 chapters/introduction.tex
Finished

    \section{Neutron star mergers} \label{sec:Intro:NeutronStarMergers}
    Binary neutron star mergers emit vast amounts of energy in gravitational waves.
    The estimated radiated energy for the GW170817 merger was $\gtrsim 5\times 10^{52}\,$ergs~\cite{GW170817Detection}, equivalent to $\sim 1\%$ of the total rest mass of the two neutron stars.
    For a source at a distance of $D$ from the observer, the amplitude of gravitational waves are  $\propto 1/D$, whereas the amplitude of electromagnetic waves are $\propto 1/D^2$.
    Additionally, electromagnetic waves can be attenuated by intervening media between the source and the detector whereas gravitational waves are transparent to the intervening media.
    Hence gravitational waves are detectable over much larger distances in comparison to electromagnetic radiation with the same source luminosity. \par
    Gravitational-wave strain emitted from the inspiral of binary neutron star mergers, approximated by Eqs.(\ref{eq:Intro:FreqEvolution}\Hyphdash*\ref{eq:Intro:chirp_ht}), can be recast in the frequency domain using the stationary phase approximation~\cite{Cutler1994}
    \begin{align}
        h_{22}(f) & \equiv \frac{\mathcal{A}}{D}\mathcal{M}^{5/6}f^{-7/6} \exp(i\Psi(f)), \label{eq:Intro:StationaryPhaseApproxH22} \\
        \Psi(f) & = 2\pi f t_c - \phi_c - \frac{\pi}{4} +\frac{3}{4}\left(8\pi\mathcal{M}f\right)^{-5/3}, \label{eq:Intro:StationaryPhaseApproxPsi} 
    \end{align}
    where $h_{22}(f)$ is the Fourier transform of the $(l,m)=(2,2)$ gravitational-wave strain, and $\Psi(f)$ is the corresponding phase.
    Here, $f$ is the frequency~(Eq.(\ref{eq:Intro:FreqEvolution})), $\phi_c$ and $t_c$ are the phase and time of coalescence, and $\mathcal{A}$ is a factor that depends on the orientation of the system.\footnote{The orientation factor, $\mathcal{A}$, depends on angles intrinsic and extrinsic to the binary system. Intrinsic angles include the angles defining the relative orientation of the neutron star spins to the orbital plane. Extrinsic angles include the angles defining the relative orientation between the orbital plane and each of the detectors.}
    \redtext{
    % Equations~\ref{eq:Intro:StationaryPhaseApproxH22}\Hyphdash*\ref{eq:Intro:StationaryPhaseApproxPsi} are valid for $a \gtrsim 6M$ where $a$ is the orbital separation of the binary system and $a\sim 6M$ is an approximation of the innermost stable circular orbit~\cite{Cutler1994}.
    Equations~(\ref{eq:Intro:StationaryPhaseApproxH22})\Hyphdash*(\ref{eq:Intro:StationaryPhaseApproxPsi}) are valid until close to the merger time.
    The orbital separation, $a$, at the merger time can be approximated by the innermost stable circular orbit, where $a \sim 6M$.
    Therefore Eqns.(\ref{eq:Intro:StationaryPhaseApproxH22})\Hyphdash*(\ref{eq:Intro:StationaryPhaseApproxPsi}) are valid for $a \gtrsim 6M$~\cite{Cutler1994}.}
    Solving Eqs.(\ref{eq:Intro:OrbitalEvolution}\Hyphdash*\ref{eq:Intro:FreqEvolution}) for $a \sim 6M$ gives $f_{max}\approx c^3 ( 6^{3/2}\pi G M )^{-1}\,\mathrm{Hz}$.
    \par

@@

    Numerical-relativity simulations are required for determining the dynamics of the late inspiral of the binary neutron star and the post-merger remnant.
    To perform a numerical-relativity simulation, the equation of state must be defined, a set of initial conditions must be evaluated in general relativity, and then Einstein's field equations must be evolved.
    Each of these three requirements make these simulations difficult.
    \redtext{The equation of state for cold neutron stars is still unknown, and the equation of state for hot post-merger remnants is even more uncertain,} after all, this is the hottest and densest bulk matter in the Universe.\par

    The initial conditions for binary neutron star mergers in general relativity are complex and a number of methods have been developed for this task.
    Codes that have been used for this include \texttt{LORENE}~\cite{Gourgoulhon2001} and \texttt{COCAL}~\cite{Tsokaros2015}.
    % The \texttt{LORENE} formulation assumes a conformally-flat approximation for  neutron star binaries and uses spectral decomposition to solve for the initial conditions~\cite{Gourgoulhon2001}.
    % This formulation constructs quasiequilibrium configurations which are approximations of the inspiral dynamics using exactly circular orbits.
    \redtext{The \texttt{LORENE} code used multi-domain spectral-methods to solve for the initial conditions of neutron star binaries assuming a conformally-flat approximation~\cite{Gourgoulhon2001}.
    This was achieved by constructing quasiequilibrium configurations, which are approximations of the inspiral dynamics using exactly circular orbits.}
    % Quasiequilibrium configurations are also generated by \texttt{COCAL}.
    % This is performed by excising regions around the neutron stars and solving Green's integral formula using multipole expansion and finite-difference methods~\cite{Uryu2012,Tsokaros2015}.
    \redtext{The \texttt{COCAL} code has also been used to generate quasiequilibrium initial conditions for binary neutron stars mergers.
    Here, the field equations are expressed in elliptic form and are solved by multipole expansion of Green's integral formula~\cite{Uryu2012,Tsokaros2015}.}
    Once the initial conditions have been approximated then the space-time can be evolved.

    Space-time evolution methods include the following formulations: Arnowitt-Deser-Misner (ADM)~\cite{Arnowitt2008}\footnote{The original book reference for ADM formalism is out of print and the authors have listed the cited paper as an `intentially unretouched' alternative.},  Baumgarte-Shapiro-Shibata-Nakamura-Oohara-Kojima (BSSNOK)~\cite{Nakamura1987,Shibata1995,Baumgarte1999,Alcubierre2000} and Z4c~\cite{Bona2003,Bona2004,Gundlach2005,Bernuzzi2010,Hilditch2013}. 
    The ADM formalism is a Hamiltonian method that splits the Einstein's field equations into separate time and space components.
    This is achieved by applying a gauge that consists of a shift vector and lapse function~\cite{Arnowitt2008}.
    % Unfortunately, ADM numerical simulations were found to be unstable to accumulated errors over the simulation time~\cite{Brandt2000,Kidder2001,Alcubierre2001a} and a number of different formulations were developed to improve this.
    \redtext{Unfortunately, as ADM numerical simulations were weakly hyperbolic in some guages, errors were found to accumulate over the simulation time~\cite{Brandt2000,Kidder2001,Alcubierre2001a}.
    To counter this, a number of different formulations were devised to improve this.}  \par

    To aid computation convergence, the BSSNOK method extended the ADM formalism by introducing a conformal factor to the three-metric and selecting a dissipative hyperbolic driver for the  shift vector~\cite{Nakamura1987,Shibata1995,Nakamura1999,Shibata1999,Baumgarte1999,Alcubierre2000,Alcubierre2001}. % Dissipative hyperbolic driver from Alcubierre2001
    The Z4c formalism modified the BSSNOK method by adding a four-vector of constraints, $Z_\mu$, to the Einstein field equations~\cite{Bona2003,Bona2004,Gundlach2005,Bernuzzi2010,Hilditch2013}. 
    This resulted in a propagating Hamiltonian constraint which helps to reduce constraint-violation growth by damping all constraint violations  except for constant modes~\cite{Gundlach2005}.
    % Constraint violations can be more easily tracked as non-vanishing values of $Z_\mu$ though particular attention needs to be paid to reflection of constraint violations at boundaries~\cite{Bernuzzi2010,Ruiz2011,Hilditch2013}.  \par
    \redtext{The introduction of $Z_\mu$, given an approapriate gauge, enforces hyperbolic solutions of the Einstein field equations by placing additional constraints on the system, thereby improving convergence~\cite{Bernuzzi2010,Ruiz2011,Hilditch2013}.} \par

@@

    The aforementioned tools allow the simulation of the late-time merger and post-merger remnant.
    However, there are a number of caveats that need to be considered.
    Firstly, the simulations are expensive, for example, 32 binary neutron star simulations took  $\sim 3\times 10^6\,$CPU hrs to complete~\cite{Takami2015}.
    Because of this, the direct use of numerical-relativity simulations in detection and parameter estimation, where we need large quantities of waveforms, is unfeasible. \par
    Secondly, numerical-relativity simulations unavoidably include numerical viscosity.
    \redtext{Early simulations show that numerical viscosity could alter the properties of both the neutron star core and the surface~\cite{Shibata2000bar,Cerda-Duran2010,Bernuzzi2012,Bernuzzi2012b}.}
    Numerical viscosity was found to introduce phase errors in the late-stage merger and improvements in numerical methods were subsequently introduced to minimise these errors~\cite{Radice2014}.
    The effects of numerical viscosity on the post-merger remnant is less clear, though it was found that numerical viscosity could alter both the neutrino luminosity~\cite{Sekiguchi2016} and the presence of one-armed spiral instability~\cite{Radice2016}.\par

@@

    Finally, simulations of the remnant are dependent on the spatial resolution.
    Generally, increasing the spatial resolution leads to more accurate simulations at the cost of simulation time.
    Ref.~\cite{Dietrich2018} state that their post-merger simulations from  the \texttt{CoRe} database are accurate enough to infer the energy and frequency content of the post-merger remnant.
    \redtext{Additionally, a recent examination of waveform systematics found that existing waveform models, including numerical-relativity simulations, were not accurate enough to inform the equation of state at very loud SNRs ($\gtrsim 100$) expected from third-generation detectors~\cite{Gamba2021}.
    These are important points: the uncertainty in the phase evolution  and collapse time of the post-merger remnant, which is discussed further in Sec.~\ref{sec:Intro:StrainAndSpectra}, has had a significant impact on the design of the models developed in this thesis.}

@@

    The timescale that sets the loss of angular momentum due to gravitational waves is given by~\cite{Paschalidis2012}
    \begin{align}
        t_{\mathrm{GW}} & \approx 150 \left(\frac{M}{2.7\,\msun}\right)^{-1}\left(\frac{R}{10\,\km}\right)^{-2}\left(\frac{f_2}{3.3\,\kHz} \right)^{-4}\left(\frac{e}{0.2} \right)^{-2}\,\mathrm{ms}, \label{eq:Intro:GWAngularMomentumLoss}        
    \end{align}
    where $f_2$ is the dominant post-merger oscillation frequency (see Sec.~\ref{sec:Intro:StrainAndSpectra}) and $e\equiv(a-b)/R$ is the ellipticity of the hypermassive neutron star. 
    Here, $a$ and $b$ are the semi-major and semi-minor axes of the neutron star, respectively, such that $R=(a+b)/2$.
    The reference value, $f_2=3.3\,\kHz$, was inferred from our model in Chapter~\ref{chapter:DetectionPE} for numerical-relativity simulation of SLy equation of state with equal-mass $1.35\,\msun$ neutron stars~\cite{Radice2016}.
    We make rough estimate for the ellipticity,  $e\approx0.2$, from the time evolution of the density profile for numerical-relativity simulation of LS220 equation of state with equal-mass $1.35\,\msun$ neutron stars~\cite{Radice2017}.
    \redtext{Numerical-relativity simulations have shown that the gravitational-wave timescale can range from $\mathcal{O}(20)\,\ms$ shortly after merger to $\gtrsim 0.5\,\s$ for longer lived simulations~\cite{Bernuzzi2016,Radice2018}.} \par

@@

    Two models of the time-domain gravitational-wave strain were used for the basis for our model in Chapter~\ref{chapter:DetectionPE}.
    The first model used a third-order exponentially-damped sinusoid and successfully matched a simulation using DD2 equation of state with equal-mass $1.35\,\msun$ neutron stars~\cite{Bauswein2016}.
    The second model implemented a quadratic frequency-drift term over time for $\fpeak$ and used this model  to measure the neutron star radius~\cite{Bose2018}.
    Spectrograms from Refs.~\cite{Rezzolla2016,Dietrich2017,Dietrich2017a} and to a lesser extent from Ref.~\cite{Maione2017}, show some frequency drifts over time in some of the frequency peaks, particularly in $\fpeak$ which lends support to the implementation of a frequency-drift term in a post-merger gravitational-wave strain model.
    \redtext{One possible cause of frequency drift would involve the contraction or expansion of the post-merger remnant, which would increase and decrease the frequency, respectively.}
      \par

@@

    A number of challenges are presented by the gravitational-wave strain generated with numerical-relativity simulations.
    Phase errors between successive resolutions accumulate over time and it is not unusual to see phase differences of $\sim 1$\,rad by the time of the merger for the finest resolutions~\cite[e.g.,][]{Radice2014b}.
    Reference~\cite{Takami2015} examined the role of the spatial resolution on the instantaneous phase of the gravitational-wave strain and found a phase difference of $0.9$\,rad between the the finest resolution tests \redtext{(See also Ref.~\cite{Breschi2019} for a discussion on post-merger waveform robustness)}.
    This results in a loss of $\sim 40\%$ of the signal-to-noise compared to a perfectly matched signal when using a matched-filter detection. \par

>>>>

    \section{Neutron star mergers} \label{sec:Intro:NeutronStarMergers}
    Binary neutron star mergers emit vast amounts of energy in gravitational waves.
    The estimated radiated energy for the GW170817 merger was $\gtrsim 5\times 10^{52}\,$ergs~\cite{GW170817Detection}, equivalent to $\sim 1\%$ of the total rest mass of the two neutron stars.
    For a source at a distance of $D$ from the observer, the amplitude of gravitational waves are  $\propto 1/D$, whereas the amplitude of electromagnetic waves are $\propto 1/D^2$.
    Additionally, electromagnetic waves can be attenuated by intervening media between the source and the detector whereas gravitational waves are transparent to the intervening media.
    Hence gravitational waves are detectable over much larger distances in comparison to electromagnetic radiation with the same source luminosity. \par
    Gravitational-wave strain emitted from the inspiral of binary neutron star mergers, approximated by Eqs.(\ref{eq:Intro:FreqEvolution}\Hyphdash*\ref{eq:Intro:chirp_ht}), can be recast in the frequency domain using the stationary phase approximation~\cite{Cutler1994}
    \begin{align}
        h_{22}(f) & \equiv \frac{\mathcal{A}}{D}\mathcal{M}^{5/6}f^{-7/6} \exp(i\Psi(f)), \label{eq:Intro:StationaryPhaseApproxH22} \\
        \Psi(f) & = 2\pi f t_c - \phi_c - \frac{\pi}{4} +\frac{3}{4}\left(8\pi\mathcal{M}f\right)^{-5/3}, \label{eq:Intro:StationaryPhaseApproxPsi} 
    \end{align}
    where $h_{22}(f)$ is the Fourier transform of the $(l,m)=(2,2)$ gravitational-wave strain, and $\Psi(f)$ is the corresponding phase.
    Here, $f$ is the frequency~(Eq.(\ref{eq:Intro:FreqEvolution})), $\phi_c$ and $t_c$ are the phase and time of coalescence, and $\mathcal{A}$ is a factor that depends on the orientation of the system.\footnote{The orientation factor, $\mathcal{A}$, depends on angles intrinsic and extrinsic to the binary system. Intrinsic angles include the angles defining the relative orientation of the neutron star spins to the orbital plane. Extrinsic angles include the angles defining the relative orientation between the orbital plane and each of the detectors.}
    Equations~(\ref{eq:Intro:StationaryPhaseApproxH22})\Hyphdash*(\ref{eq:Intro:StationaryPhaseApproxPsi}) are valid until close to the merger time.
    The orbital separation, $a$, at the merger time can be approximated by the innermost stable circular orbit, where $a \sim 6M$.
    Therefore Eqns.(\ref{eq:Intro:StationaryPhaseApproxH22})\Hyphdash*(\ref{eq:Intro:StationaryPhaseApproxPsi}) are valid for $a \gtrsim 6M$~\cite{Cutler1994}.
    Solving Eqs.(\ref{eq:Intro:OrbitalEvolution}\Hyphdash*\ref{eq:Intro:FreqEvolution}) for $a \sim 6M$ gives $f_{max}\approx c^3 ( 6^{3/2}\pi G M )^{-1}\,\mathrm{Hz}$.
    \par

@@

    Numerical-relativity simulations are required for determining the dynamics of the late inspiral of the binary neutron star and the post-merger remnant.
    To perform a numerical-relativity simulation, the equation of state must be defined, a set of initial conditions must be evaluated in general relativity, and then Einstein's field equations must be evolved.
    Each of these three requirements make these simulations difficult.
    The equation of state for cold neutron stars is still unknown, and the equation of state for hot post-merger remnants is even more uncertain, after all, this is the hottest and densest bulk matter in the Universe.\par

    The initial conditions for binary neutron star mergers in general relativity are complex and a number of methods have been developed for this task.
    Codes that have been used for this include \texttt{LORENE}~\cite{Gourgoulhon2001} and \texttt{COCAL}~\cite{Tsokaros2015}.
    The \texttt{LORENE} code used multi-domain spectral-methods to solve for the initial conditions of neutron star binaries assuming a conformally-flat approximation~\cite{Gourgoulhon2001}.
    This was achieved by constructing quasiequilibrium configurations, which are approximations of the inspiral dynamics using exactly circular orbits.
    The \texttt{COCAL} code has also been used to generate quasiequilibrium initial conditions for binary neutron stars mergers.
    Here, the field equations are expressed in elliptic form and are solved by multipole expansion of Green's integral formula~\cite{Uryu2012,Tsokaros2015}.
    Once the initial conditions have been approximated then the space-time can be evolved.

    Space-time evolution methods include the following formulations: Arnowitt-Deser-Misner (ADM)~\cite{Arnowitt2008}\footnote{The original book reference for ADM formalism is out of print and the authors have listed the cited paper as an `intentially unretouched' alternative.},  Baumgarte-Shapiro-Shibata-Nakamura-Oohara-Kojima (BSSNOK)~\cite{Nakamura1987,Shibata1995,Baumgarte1999,Alcubierre2000} and Z4c~\cite{Bona2003,Bona2004,Gundlach2005,Bernuzzi2010,Hilditch2013}. 
    The ADM formalism is a Hamiltonian method that splits the Einstein's field equations into separate time and space components.
    This is achieved by applying a gauge that consists of a shift vector and lapse function~\cite{Arnowitt2008}.
    Unfortunately, as ADM numerical simulations are weakly hyperbolic in some guages, errors were found to accumulate over the simulation time~\cite{Brandt2000,Kidder2001,Alcubierre2001a}.
    To counter this, a number of different formulations were devised to improve this.  \par

    To aid computation convergence, the BSSNOK method extended the ADM formalism by introducing a conformal factor to the three-metric and selecting a dissipative hyperbolic driver for the  shift vector~\cite{Nakamura1987,Shibata1995,Nakamura1999,Shibata1999,Baumgarte1999,Alcubierre2000,Alcubierre2001}. % Dissipative hyperbolic driver from Alcubierre2001
    The Z4c formalism modified the BSSNOK method by adding a four-vector of constraints, $Z_\mu$, to the Einstein field equations~\cite{Bona2003,Bona2004,Gundlach2005,Bernuzzi2010,Hilditch2013}. 
    This resulted in a propagating Hamiltonian constraint which helps to reduce constraint-violation growth by damping all constraint violations  except for constant modes~\cite{Gundlach2005}.
    The introduction of $Z_\mu$, given an approapriate gauge, enforces hyperbolic solutions of the Einstein field equations by placing additional constraints on the system, thereby improving convergence~\cite{Bernuzzi2010,Ruiz2011,Hilditch2013}. \par

@@

    The aforementioned tools allow the simulation of the late-time merger and post-merger remnant.
    However, there are a number of caveats that need to be considered.
    Firstly, the simulations are expensive, for example, 32 binary neutron star simulations took  $\sim 3\times 10^6\,$CPU hrs to complete~\cite{Takami2015}.
    Because of this, the direct use of numerical-relativity simulations in detection and parameter estimation, where we need large quantities of waveforms, is unfeasible. \par
    Secondly, numerical-relativity simulations unavoidably include numerical viscosity.
    Early simulations show that numerical viscosity could alter the properties of both the neutron star core and the surface~\cite{Shibata2000bar,Cerda-Duran2010,Bernuzzi2012,Bernuzzi2012b}.
    Numerical viscosity was found to introduce phase errors in the late-stage merger and improvements in numerical methods were subsequently introduced to minimise these errors~\cite{Radice2014}.
    The effects of numerical viscosity on the post-merger remnant is less clear, though it was found that numerical viscosity could alter both the neutrino luminosity~\cite{Sekiguchi2016} and the presence of one-armed spiral instability~\cite{Radice2016}.\par

@@

    Finally, simulations of the remnant are dependent on the spatial resolution.
    Generally, increasing the spatial resolution leads to more accurate simulations at the cost of simulation time.
    Ref.~\cite{Dietrich2018} state that their post-merger simulations from  the \texttt{CoRe} database are accurate enough to infer the energy and frequency content of the post-merger remnant.
    Additionally, a recent examination of waveform systematics found that existing waveform models, including numerical-relativity simulations, were not accurate enough to inform the equation of state at very loud SNRs ($\gtrsim 100$) expected from third-generation detectors~\cite{Gamba2021}.
    These are important points: the uncertainty in the phase evolution  and collapse time of the post-merger remnant, which is discussed further in Sec.~\ref{sec:Intro:StrainAndSpectra}, has had a significant impact on the design of the models developed in this thesis.

@@

    The timescale that sets the loss of angular momentum due to gravitational waves is given by~\cite{Paschalidis2012}
    \begin{align}
        t_{\mathrm{GW}} & \approx 150 \left(\frac{M}{2.7\,\msun}\right)^{-1}\left(\frac{R}{10\,\km}\right)^{-2}\left(\frac{f_2}{3.3\,\kHz} \right)^{-4}\left(\frac{e}{0.2} \right)^{-2}\,\mathrm{ms}, \label{eq:Intro:GWAngularMomentumLoss}        
    \end{align}
    where $f_2$ is the dominant post-merger oscillation frequency (see Sec.~\ref{sec:Intro:StrainAndSpectra}) and $e\equiv(a-b)/R$ is the ellipticity of the hypermassive neutron star. 
    Here, $a$ and $b$ are the semi-major and semi-minor axes of the neutron star, respectively, such that $R=(a+b)/2$.
    The reference value, $f_2=3.3\,\kHz$, was inferred from our model in Chapter~\ref{chapter:DetectionPE} for numerical-relativity simulation of SLy equation of state with equal-mass $1.35\,\msun$ neutron stars~\cite{Radice2016}.
    We make rough estimate for the ellipticity,  $e\approx0.2$, from the time evolution of the density profile for numerical-relativity simulation of LS220 equation of state with equal-mass $1.35\,\msun$ neutron stars~\cite{Radice2017}.
    Numerical-relativity simulations have shown that the gravitational-wave timescale can range from $\mathcal{O}(20)\,\ms$ shortly after merger to $\gtrsim 0.5\,\s$ for longer lived simulations~\cite{Bernuzzi2016,Radice2018}. \par

@@

    Two models of the time-domain gravitational-wave strain were used for the basis for our model in Chapter~\ref{chapter:DetectionPE}.
    The first model used a third-order exponentially-damped sinusoid and successfully matched a simulation using DD2 equation of state with equal-mass $1.35\,\msun$ neutron stars~\cite{Bauswein2016}.
    The second model implemented a quadratic frequency-drift term over time for $\fpeak$ and used this model  to measure the neutron star radius~\cite{Bose2018}.
    Spectrograms from Refs.~\cite{Rezzolla2016,Dietrich2017,Dietrich2017a} and to a lesser extent from Ref.~\cite{Maione2017}, show some frequency drifts over time in some of the frequency peaks, particularly in $\fpeak$ which lends support to the implementation of a frequency-drift term in a post-merger gravitational-wave strain model.
    One possible cause of frequency drift would involve the contraction or expansion of the post-merger remnant, which would increase and decrease the frequency, respectively.
      \par

@@

    A number of challenges are presented by the gravitational-wave strain generated with numerical-relativity simulations.
    Phase errors between successive resolutions accumulate over time and it is not unusual to see phase differences of $\sim 1$\,rad by the time of the merger for the finest resolutions~\cite[e.g.,][]{Radice2014b}.
    Reference~\cite{Takami2015} examined the role of the spatial resolution on the instantaneous phase of the gravitational-wave strain and found a phase difference of $0.9$\,rad between the the finest resolution tests (See also Ref.~\cite{Breschi2019} for a discussion on post-merger waveform robustness).
    This results in a loss of $\sim 40\%$ of the signal-to-noise compared to a perfectly matched signal when using a matched-filter detection. \par

### 2021-10-17 papers/paper1.tex
Removed unused comments
Finished

% !TeX root = ../Thesis.tex
\documentclass[../Thesis.tex]{subfiles}

% \documentclass[
% % preprint,
% % prx,
% % prl,
% % twocolumn,
% %onecolumn,%!TEX encoding = UTF-8 Unicode
% nofootinbib,
% % ,secularistic%
% tightenlines %
% floatfix,
% showpacs ,amssymb, aps,% linenumbers
% superscriptaddress
% ]{revtex4-1}
% \documentclass[../main.tex]{subfiles}
% \usepackage{amsmath,slashed}
% \usepackage[left]{lineno}

% \usepackage{comment}
% \usepackage{graphicx}
% \usepackage{amsmath,amssymb}
% \usepackage{bm}
% \usepackage[shortlabels]{enumitem}
% \usepackage{float,soul}
% \usepackage{natbib}
% \usepackage{wasysym}
% \usepackage{aas_macros}
% \graphicspath{{images/}}
% \RequirePackage{lineno}
% \usepackage[usenames,dvipsnames]{color}

% % \newcommand{\red}[1]{\textcolor{red}{#1}}
% % \newcommand{\pdl}[1]{\textcolor{red}{[PDL: #1]}}
% % \newcommand{\pje}[1]{\textcolor{Purple}{[PJE: #1]}}
% % \newcommand{\blue}[1]{\textcolor{blue}{#1}}
% \DeclareMathOperator*{\argmin}{arg\,min}
% \newcommand{\SPA}{School of Physics and Astronomy, Monash University, Vic 3800, Australia}
% \newcommand{\OzGravMonash}{OzGrav: The ARC Centre of Excellence for Gravitational Wave Discovery, Clayton VIC 3800, Australia}

@@

\chapterauthor{Paul J. Easter \\ Paul D. Lasky\\Andrew R. Casey\\Luciano Rezzolla\\Kentaro Takami}
% \email{paul.easter@monash.edu}
% \author{Paul D. Lasky}
% \email{paul.lasky@monash.edu}
% \affiliation{\SPA}
% \affiliation{\OzGravMonash}
% \author{Andrew R. Casey}
% \email{andrew.casey@monash.edu}
% \affiliation{\SPA}
% \affiliation{\OzGravMonash}
% \affiliation{Faculty of Information Technology, Monash University, VIC 3800, Australia}
% \author{Luciano Rezzolla}
% \affiliation{Institut f\"ur Theoretische Physik, Max-von-Laue-Stra\ss e 1, 60438 Frankfurt, Germany}
% \author{Kentaro Takami}
% \affiliation{Kobe City College of Technology, 651-2194 Kobe, Japan}

@@

        It is possible that a phase transition may occur in the remnant during the post-merger phase due to an increase in density~(e.g.~\cite{Most2018,Bauswein2018}). This could be detected by comparing the inferred values of the quadrupolar tidal coupling constant from the inspiral and post-merger phases. Simulations that compare hydrodynamic models with and without quark deconfinement show that the presence of quark deconfinement causes no change in the inspiral gravitational-wave signal due to the low densities involved. This is contrasted to the post-merger phase where the densities are greater, leading to a higher proportion of deconfined quarks when modelled, which in turn leads to a change in the post-merger gravitational-wave strain~\cite{Most2018b}. Phase transitions due to quark deconfinement could be detected by training our model on simulations that disallow quark deconfinement. If a post-merger signal was detected that contained the signature of quark-deconfinement then this would result in an inconsistency between the tidal coupling constant inferred from the post-merger gravitational-wave spectra and the inspiral phase. We note that other mechanisms, either physical or arising from errors in the modelling, could also cause such frequency shifts. We leave the exploration of this as future work. \par

@@

        Results based on our trained model suggest that the model is selective and could potentially be used in parameter estimation of detected events. If posteriors for the mass and tidal coupling constant are able to be determined, then it is a simple step to calculate the posteriors for the compactness, Eq.(\ref{eq:Ccalc1}), and the radius of the neutron star. This will be confirmed in future work using a Bayesian framework. Parameter estimation of the post-merger spectra could set bounds on the post-merger quadrupolar tidal coupling constant, allowing comparison with the inspiral value. This could determine whether a phase change in the equation of state is present~\cite{Bauswein2018,Most2018b}.\par

>>>>

%!TeX root = ../Thesis.tex
\documentclass[../Thesis.tex]{subfiles}

@@

\chapterauthor{Paul J. Easter \\ Paul D. Lasky\\Andrew R. Casey\\Luciano Rezzolla\\Kentaro Takami}

@@

        It is possible that a phase transition may occur in the remnant during the post-merger phase due to an increase in density~(e.g.~\cite{Most2018,Bauswein2019}). This could be detected by comparing the inferred values of the quadrupolar tidal coupling constant from the inspiral and post-merger phases. Simulations that compare hydrodynamic models with and without quark deconfinement show that the presence of quark deconfinement causes no change in the inspiral gravitational-wave signal due to the low densities involved. This is contrasted to the post-merger phase where the densities are greater, leading to a higher proportion of deconfined quarks when modelled, which in turn leads to a change in the post-merger gravitational-wave strain~\cite{Most2018b}. Phase transitions due to quark deconfinement could be detected by training our model on simulations that disallow quark deconfinement. If a post-merger signal was detected that contained the signature of quark-deconfinement then this would result in an inconsistency between the tidal coupling constant inferred from the post-merger gravitational-wave spectra and the inspiral phase. We note that other mechanisms, either physical or arising from errors in the modelling, could also cause such frequency shifts. We leave the exploration of this as future work. \par

@@

        Results based on our trained model suggest that the model is selective and could potentially be used in parameter estimation of detected events. If posteriors for the mass and tidal coupling constant are able to be determined, then it is a simple step to calculate the posteriors for the compactness, Eq.(\ref{eq:Ccalc1}), and the radius of the neutron star. This will be confirmed in future work using a Bayesian framework. Parameter estimation of the post-merger spectra could set bounds on the post-merger quadrupolar tidal coupling constant, allowing comparison with the inspiral value. This could determine whether a phase change in the equation of state is present~\cite{Bauswein2019,Most2018b}.\par
